Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-09-07 tex/merf/TALMajor/gui.tex
applied recommended edits; almost done

\subsection{Tag type Boolean formula editor}
The user writes MBF tag types with the tag type editor introduced in~\cite{JaZaMatar}. %shown in Figure~\ref{f:bfe}.
First the user specifies atomic terms by selecting a feature from ${\cal F}$. 
%The pattern filters the feature values.
The user can also choose whether to require an exact match using the
\cci{isA} predicate, or a substring match using the
\cci{contains} predicate option.

@@

\subsection{Tag type regular expression editor}
\vspace{-1em}

After interacting with the MBF editor, the user moves to 
specify the regular expressions. 
The MRE editor of Figure~\ref{f:sfe} allows the 
definition of an MRE tag type in a user friendly manner. 
The user first adds the required MBF formulae 
by selecting a label from ${\cal T}$ under MBFs. 
The Boolean formula of a highlighted tag type is shown in the table on the lower left pane. 
Each selected MBF is associated with an automatic name. 
The user can nest the MRE expression using a tree view of the MRE operations. 
The tree features the name, MBF, and operation for each sub-expression. 
%Moreover, a defined MRE is directly added to the MBF list enabling substitution and recursion 
%in formula definition.

@@

%The regular expression editor enables the user to apply operations to the selected expressions. 
To specify a binary operation the user selects two sub-expressions and clicks the corresponding
operation button. 
%To do so, the user selects one or two expressions then selects an operation from the ones shown 
%in the lower left of the view. 
The operations include disjunction, conjunction, zero or one, sequence, zero or more, 
one or more, and up to a user defined constant.
The right pane shows a description of the tag type and a set of legend 
descriptors.

%\[
%	(P|N)+~\mathit{O}?~R~\mathit{O^{\wedge}2}~(P|N|U)+
%\]
%In the direction task, we aim to build the expression shown above. 
%Thus we start by adding the MBFs $P$, $N$, $\textit{NONE}$, $R$, $\textit{NONE}$, $P$, $N$, and $U$. 
%We apply the required operations incrementally to build the final expression. 
%For example, we apply the disjunction ($|$) operation to the first $P$ and $N$ added, 
%then we apply the one or more ($+$) operation to the resulting sub-expression.
%We proceed as such to build the regular expression shown in Figure~\ref{f:sfe}. 
%Similar to the MBF tag type editor, the user specifies a set of legend descriptors for each MRE tag type.
%

@@

While specifying an MRE the user can interact with the visualization and editor views
to make sure the MRE expresses the intent. 
The color-sensitive text view in Figure~\ref{fig:treegraph} shows 
the highlighted tag matches after the user called the MRE simulator using 
the {\tt Tagtypes} menu.

The matching parse tree view shows the selected match in a graph view.
Figure~\ref{fig:treegraph} shows the matching parse tree of the direction task 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(Dubai Mall is located near this building). 
%Leaf nodes show the match text, internal nodes show the operations, and edge labels shows the MBF tag type name.

@@

After the user is satisfied with the MRE matches, 
the user moves to define relations and code actions. 
The relation editor allows the user to define relations 
by specifying $\langle e_1,e_2,r\rangle$ tuples, 
where $e_1$ and $e_2$ denote source and destination entities, and $r$ denotes 
the label.
The editor shows the MRE tree and allows the user to select the sub 
expressions and select features 
of the matches of the sub expressions to define the three components of the relation.

A snapshot of the GUI in 
Figure~\ref{fig:treegraph} shows in an interactive graph view
the entity-relation graph of the match of the user defined relation 
extracted from the matching parse tree of the MRE. 
%
In the computational action editor, an advanced user can 
enter C++ code and use the \framework API to program and process 
subexpression matches. 
%RE matches In order to add computational actions to an MRE sub-expression, 
%the user selects to edit the actions in the MRE tag type editor. 
%The view allows the user to specify the actions in a user friendly manner. 
%\framework provides the user with an API to access the sub-expression match features easily and define the pre-match actions. 
%The match features include the text, position, length, and numerical value if applicable, and morphological features for MBF matches. 
%\framework also enables the user to add C++ declarations and library includes.

>>>>

\subsection{Tag type Boolean formula editor}
The user writes MBF tag types with the tag type editor introduced in~\citename{JaZaMatar}~\citeyear{JaZaMatar}. %shown in Figure~\ref{f:bfe}.
First the user specifies atomic terms by selecting a feature from ${\cal F}$. 
%The pattern filters the feature values.
The user can also choose whether to require an exact match using the
\cci{isA} predicate, or a substring match using the
\cci{contains} predicate option.

@@

\subsection{Tag type regular expression editor}
%\vspace{-1em}

After interacting with the MBF editor, the user moves to 
specify the regular expressions. 
The MRE editor of Figure~\ref{f:sfe} allows the 
definition of an MRE tag type in a user-friendly manner. 
The user first adds the required MBF formulae 
by selecting a label from ${\cal T}$ under MBFs. 
The Boolean formula of a highlighted tag type is shown in the table on the lower left pane. 
Each selected MBF is associated with an automatic name. 
The user can nest the MRE expression using a tree view of the MRE operations. 
The tree features the name, MBF, and operation for each subexpression. 
%Moreover, a defined MRE is directly added to the MBF list enabling substitution and recursion 
%in formula definition.

@@

%The regular expression editor enables the user to apply operations to the selected expressions. 
To specify a binary operation the user selects two subexpressions and clicks the corresponding
operation button. 
%To do so, the user selects one or two expressions then selects an operation from the ones shown 
%in the lower left of the view. 
The operations include disjunction, conjunction, zero or one, sequence, zero or more, 
one or more, and up to a user defined constant.
The right pane shows a description of the tag type and a set of legend 
descriptors.

%\[
%	(P|N)+~\mathit{O}?~R~\mathit{O^{\wedge}2}~(P|N|U)+
%\]
%In the direction task, we aim to build the expression shown above. 
%Thus we start by adding the MBFs $P$, $N$, $\textit{NONE}$, $R$, $\textit{NONE}$, $P$, $N$, and $U$. 
%We apply the required operations incrementally to build the final expression. 
%For example, we apply the disjunction ($|$) operation to the first $P$ and $N$ added, 
%then we apply the one or more ($+$) operation to the resulting subexpression.
%We proceed as such to build the regular expression shown in Figure~\ref{f:sfe}. 
%Similar to the MBF tag type editor, the user specifies a set of legend descriptors for each MRE tag type.
%

@@

While specifying an MRE, the user can interact with the visualization and editor views
to make sure the MRE expresses the intent. 
The color-sensitive text view in Figure~\ref{fig:treegraph} shows 
the highlighted tag matches after the user called the MRE simulator using 
the {\tt Tagtypes} menu.

The matching parse tree view shows the selected match in a graph view.
Figure~\ref{fig:treegraph} shows the matching parse tree of the direction task 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(``Dubai Mall is located near this building''). 
%Leaf nodes show the match text, internal nodes show the operations, and edge labels shows the MBF tag type name.

@@

After the user is satisfied with the MRE matches, 
the user moves to define relations and code actions. 
The relation editor allows the user to define relations 
by specifying $\langle e_1,e_2,r\rangle$ tuples, 
where $e_1$ and $e_2$ denote source and destination entities, and $r$ denotes 
the label.
The editor shows the MRE tree and allows the user to select the subexpressions and select features 
of the matches of the subexpressions to define the three components of the relation.

A snapshot of the GUI in 
Figure~\ref{fig:treegraph} shows in an interactive graph view
the entity-relation graph of the match of the user defined relation 
extracted from the matching parse tree of the MRE. 
%
In the computational action editor, an advanced user can 
enter C++ code and use the \framework API to program and process 
subexpression matches. 
%RE matches In order to add computational actions to an MRE subexpression, 
%the user selects to edit the actions in the MRE tag type editor. 
%The view allows the user to specify the actions in a user friendly manner. 
%\framework provides the user with an API to access the subexpression match features easily and define the pre-match actions. 
%The match features include the text, position, length, and numerical value if applicable, and morphological features for MBF matches. 
%\framework also enables the user to add C++ declarations and library includes.

### 2018-09-07 tex/merf/TALMajor/introduction.tex
final version
applied recommended edits; almost done

Figure~\ref{fig:intromotiv} illustrates the target of \framework using 
the directions to Dubai Mall example~\footnote{text taken from the Dubai 
Mall website ~\url{http://www.thedubaimall.com/ar/}}.
The figure also presents a transliteration and an English translation of 
the Arabic text.
The framed words in the text are entities referring to names of people 
($n_1,n_2,n_3$), 
names of places ($p_1,\dots,p_7$), 
relative positions ($r_1,\dots,r_4$), 
and numerical terms ($u_1,u_2$). 
We would like to extract those entities, and then extract
the relational entities forming the graph in Figure~\ref{fig:intromotiv} 
where vertices express entities, 
and edges represent the relational entities.

@@

An editor allows the user to associate an action with each sub-expression.
The user specifies the action with C++ code and uses an API to access information 
related to the matches such as text, position, length, morphological features, 
and numerical value.
%
Each regular expression is associated with a named identifier to form a 
{\em local grammar} like structure~\cite{traboulsi2009arabic}. 
%
A relation definition GUI allows the user to provide relational tuples 
where each tuple has a 
source, a destination and an edge label.
The user uses the regular expression identifiers to define the relational 
tuple elements.
For example, the relation between $e_1, e_2$ and $r$ shown in 
Figure~\ref{fig:intromotiv} is a match of a relational tuple 
over the components of $re$. 
%
We refer to regular expressions and Boolean formulae as expressions and formulae, respectively.
We also refer to expressions as rules
when used in a grammar context; e.g.  when used with an identifier. 
%We discuss the importance of the morphological features supported in 
%\framework terms in Section~\ref{sec:morph}; in brief, morphological preprocessing
%is key to Arabic NLP.% due to the richness of Arabic morphology.

@@

This work significantly extends~\cite{JaZaMatar} 
that allows for manual, and morphology annotation.
\framework enables a user to incrementally create 
complex annotations for Arabic based on automatic 
extraction of morphological tags through a user friendly interactive interface. 
\framework has the following advantages.
\begin{itemize}
  \item \framework provides a novel and intuitive visual interface to build formulae over morphological features, 
    build regular expressions over the resulting formulae, and thereafter compute automatic tags.
  \item To our knowledge, this morphology-based framework is the first for Arabic entity and relational entity extraction.
  \item \framework provides the user with the ability to rapidly create annotated Arabic text corpora with sophisticated morphology-based tags.
\end{itemize}

In \framework, we make the following contributions.
\begin{itemize}
  \item \framework enables the user to define relations in a simple manner
    and automatically detects relational entities matching the user defined relations. 
  \item \framework enables the user to associate sub-expressions
    with code actions, and executes the code action 
    when a corresponding match is found.
    It also provides an API to enable access to match 
    features such as text, position, length, numerical value, and morphological features.
  \item \framework enables the user to tag words based on a novel light Arabic wordnet relation that leverages the synonym $Syn^k$ feature.
  \item \framework is open source and available online for the research community 
    under on (\url{https://github.com/codelogicanalysis/atmine}).
\end{itemize}

>>>>

Figure~\ref{fig:intromotiv} illustrates the target of \framework using 
the directions to Dubai Mall example~\footnote{Text taken from the Dubai 
Mall website ~\url{http://www.thedubaimall.com/ar/}.}.
The figure also presents a transliteration and an English translation of 
the Arabic text.
The framed words in the text are entities referring to names of people 
($n_1,n_2,n_3$), 
names of places ($p_1,\dots,p_7$), 
relative positions ($r_1,\dots,r_4$), 
and numerical terms ($u_1,u_2$). 
We would like to extract those entities, and then extract
the relational entities forming the graph in Figure~\ref{fig:intromotiv} 
where vertices express entities, 
and edges represent the relational entities.

@@

An editor allows the user to associate an action with each subexpression.
The user specifies the action with C++ code and uses an API to access information 
related to the matches such as text, position, length, morphological features, 
and numerical value.
%
Each regular expression is associated with a named identifier to form a 
{\em local grammar} like structure~\cite{traboulsi2009arabic}. 
%
A relation definition GUI allows the user to provide relational tuples 
where each tuple has a 
source, a destination and an edge label.
The user uses the regular expression identifiers to define the relational 
tuple elements.
For example, the relation between $e_1, e_2$ and $r$ shown in 
Figure~\ref{fig:intromotiv} is a match of a relational tuple 
over the components of $re$. 
%
We refer to regular expressions and Boolean formulae as expressions and formulae, respectively.
We also refer to expressions as rules
when used in a grammar context; e.g.  when used with an identifier. 
%We discuss the importance of the morphological features supported in 
%\framework terms in Section~\ref{sec:morph}; in brief, morphological preprocessing
%is key to Arabic NLP.% due to the richness of Arabic morphology.

@@

This work significantly extends~\citename{JaZaMatar}~\citeyear{JaZaMatar}
that allows for manual, and morphology annotation.
\framework enables a user to incrementally create 
complex annotations for Arabic based on automatic 
extraction of morphological tags through a user-friendly interactive interface. 
\framework has the following advantages.
\begin{itemize}
  \item \framework provides a novel and intuitive visual interface to build formulae over morphological features, 
    build regular expressions over the resulting formulae, and thereafter compute automatic tags.
  \item To our knowledge, this morphology-based framework is the first for Arabic entity and relational entity extraction.
  \item \framework provides the user with the ability to rapidly create annotated Arabic text corpora with sophisticated morphology-based tags.
\end{itemize}

In \framework, we make the following contributions.
\begin{itemize}
  \item \framework enables the user to define relations in a simple manner
    and automatically detects relational entities matching the user defined relations. 
  \item \framework enables the user to associate subexpressions
    with code actions, and executes the code action 
    when a corresponding match is found.
    It also provides an API to enable access to match 
    features such as text, position, length, numerical value, and morphological features.
  \item \framework enables the user to tag words based on a novel light Arabic WordNet relation that leverages the synonym $Syn^k$ feature.
  \item \framework is open source and available online for the research community 
    under \url{https://github.com/codelogicanalysis/atmine}.
\end{itemize}

### 2018-09-07 tex/merf/TALMajor/relatedwork.tex
applied recommended edits; almost done

TEXTMARKER is a semi-automatic rule-based IE system 
for structured data acquisition\cite{atzmueller2008rule}.
%TEXTMARKER requires user defined rules that consider features of the text 
%However, constructing the rules requires manual work of a domain specialist. 
Both TEXTMARKER and \framework provide the user with GUI editor and result visualizer.
%\framework also differs in that it provides graph visualization for regular expression and user defisemantic relation matches.

The work in~\cite{urbain2012user} presents a user 
driven relational model and targets entity and relation extraction. 
The user enters a natural language query, and uses the OpenNLP toolkit to 
extract tags and relations from the query. 
%Then, it extends the extracted entities into a query model, 
%and uses the query model to extract matches from the document. 
Similar to \framework, the system constructs entities and relations.

@@

\begin{table}[tb]
\resizebox{\columnwidth}{!}{
\begin{tabular}{l|c|c|c|c|c}
%{p{2.7cm}|p{.7cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}}
Features & \framework & SystemT & TEXT\linebreak MARKER & Urbain & QARAB\\ \hline
Query type & MRE & AQL & matching \linebreak rules & natural language & natural language \\ \hline
Morphology support& $\checked$ & - & - & OpenNLP & Parser \\ \hline
Relations & $\checked$ & - & - & $\checked$ & - \\ \hline
Actions & $\checked$ & - & - & - & - \\ \hline
Editor & $\checked$ & - & $\checked$ & - & - \\ \hline
Tag visualization & $\checked$ & - & $\checked$ & - & - \\ \hline
Graph visualization & $\checked$ & - & - & - & - \\
\end{tabular}
}
\caption{\label{tab:iecomp}Comparison of \framework with SystemT, 
TEXTMARKER, 
Urbain, 
QARAB}
\end{table}

Table~\ref{tab:iecomp} summarizes the comparison between \framework and
other systems. \framework differs in that it provides 
code actions, user defined relations, and an interactive 
graph visualization of the relational entities. 
It also differs in that it fully supports Arabic morphological analysis
while only QARAB supports Arabic linguistic features using a parser, and 
the work in ~\cite{urbain2012user} uses OpenNLP that currently lacks full
support for Arabic morphological features. 
Similar to TEXTMARKER, \framework has the advantage of providing 
a user friendly interactive interface to edit the entity and relational 
specifications and visualize the results.

@@

Another track in the literature targets specific tasks such as NER using 
statistical and machine learning techniques such as maximum entropy, 
optimized feature sets and conditional random fields~\cite{benajiba2007anersys,benajiba2008arabic,ekbal2010named,abdelrahman2010integrated}.
%The work in~\cite{ekbal2010named} presents a language independent approach for NER extraction using {\em support vector machines}. 
%The work in~\cite{abdelrahman2010integrated} integrates a semi-supervised bootstrapping pattern recognition technique, and a supervised classifier based on {\em conditional random fields} to solve NER problems.
Knowledge-based techniques such 
as~\cite{zaghouani2010adapting,traboulsi2009arabic} propose local grammars with morphological stemming.% to perform NER. 
%The work in~
\cite{ZaMaHaCicling2012Entity} present a method for extracting entities and events, and relations among them from Arabic text using a hierarchy of manually built 
finite state machines driven by morphological features, 
and graph transformation algorithms. 
Such techniques require advanced linguistic and programming expertise.

@@

{\bf Annotation tools.}
MMAX2 is a manual multi-level linguistic annotation tool with an XML 
based data model~\cite{mmax2}.
%It enables the user to create, browse, visualize, and query annotations and may be able to resolve coreference tags. 
BRAT~\cite{brat} and WordFreak~\cite{wordfreak} are manual 
multi-lingual user friendly web-based annotators that allow the construction 
of entity and relation annotation corpora.
%They can be extended through plug-ins to enable automatic annotators and customized annotation visualization specifications.
Knowtator~\cite{ogren2006knowtator} is a general purpose incremental text annotation tool 
implemented as a Prot\'eg\'e~\cite{gennari2003evolution} plug-in. 
Prot\'eg\'e is an open-source platform with a suite of tools to construct domain models and knowledge-based applications with ontology. 
However, it doesn't support the Arabic language.

@@

\cite{kholidy2010towards} present an overview of annotation tools 
%and their Arabic-English word alignment issues 
and concludes with a set of rules and guidelines needed in an Arabic annotation alignment tool.
The work in \cite{dukes2011supervised} presents a collaborative effort towards morphological and syntactic annotation of the Quran.
\cite{dorr2010interlingual} presents a framework for inter-lingual 
annotation of parallel text corpora with multi-level representations.
\cite{kulick2010consistent} presents the integration of the 
Standard Arabic Morphological Analyzer (SAMA) into the workflow of the Arabic Treebank.

The work in \cite{smrz2004morphotrees} presents %the extension of TrEd, 
a customizable general purpose tree editor, with the Arabic MorphoTrees annotations. 
The MorphoTrees present the morphological analyses in a hierarchical organization based on common features.

Task specific annotation tools such as~\cite{alrahabi2006semantic}
use enunciation semantic maps to automatically annotate 
directly reported Arabic and French speech.
AraTation is another task specific tool for semantic annotation of 
Arabic news using web ontology based semantic maps~\cite{saleh2009aratation}.
We differ in that \framework is general, and not task specific, and it uses 
morphology-based features as atomic terms.
Fassieh is a commercial Arabic text annotation tool that enables the production of large Arabic text corpora~\cite{attia2009fassieh}.
The tool supports Arabic text factorization including morphological analysis, POS tagging, 
full phonetic transcription, and lexical semantics analysis in an automatic mode.
Fassieh is not directly accessible to the research community and requires commercial licensing.
\framework is open source and differs in that it allows the user to build tag types and extract entities and relations from text.

>>>>

TEXTMARKER is a semi-automatic rule-based IE system 
for structured data acquisition~\cite{atzmueller2008rule}.
%TEXTMARKER requires user defined rules that consider features of the text 
%However, constructing the rules requires manual work of a domain specialist. 
Both TEXTMARKER and \framework provide the user with GUI editor and result visualizer.
%\framework also differs in that it provides graph visualization for regular expression and user defisemantic relation matches.

The work in~\citename{urbain2012user}~\citeyear{urbain2012user} presents a user-driven relational model and targets entity and relation extraction. 
The user enters a natural language query, and uses the OpenNLP toolkit to 
extract tags and relations from the query. 
%Then, it extends the extracted entities into a query model, 
%and uses the query model to extract matches from the document. 
Similar to \framework, the system constructs entities and relations.

@@

\begin{table}[tb]
\resizebox{\columnwidth}{!}{
\begin{tabular}{l|c|c|c|c|c}
%{p{2.7cm}|p{.7cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}}
Features & \framework & SystemT & TEXT\linebreak MARKER & Urbain & QARAB\\ \hline
Query type & MRE & AQL & matching \linebreak rules & natural language & natural language \\ \hline
Morphology support& $\checked$ & - & - & OpenNLP & Parser \\ \hline
Relations & $\checked$ & - & - & $\checked$ & - \\ \hline
Actions & $\checked$ & - & - & - & - \\ \hline
Editor & $\checked$ & - & $\checked$ & - & - \\ \hline
Tag visualization & $\checked$ & - & $\checked$ & - & - \\ \hline
Graph visualization & $\checked$ & - & - & - & - \\
\end{tabular}
}
\caption{\label{tab:iecomp}Comparison of \framework with SystemT, 
TEXTMARKER, 
Urbain, 
QARAB.}
\end{table}

Table~\ref{tab:iecomp} summarizes the comparison between \framework and
other systems. \framework differs in that it provides 
code actions, user defined relations, and an interactive 
graph visualization of the relational entities. 
It also differs in that it fully supports Arabic morphological analysis
while only QARAB supports Arabic linguistic features using a parser, and 
the work in~\citename{urbain2012user}~\citeyear{urbain2012user} uses OpenNLP that currently lacks full
support for Arabic morphological features. 
Similar to TEXTMARKER, \framework has the advantage of providing 
a user-friendly interactive interface to edit the entity and relational 
specifications and visualize the results.

@@

Another track in the literature targets specific tasks such as NER using 
statistical and machine-learning techniques such as maximum entropy, 
optimized feature sets and conditional random fields~\cite{benajiba2007anersys,benajiba2008arabic,ekbal2010named,abdelrahman2010integrated}.
%The work in~\cite{ekbal2010named} presents a language independent approach for NER extraction using {\em support vector machines}. 
%The work in~\cite{abdelrahman2010integrated} integrates a semi-supervised bootstrapping pattern recognition technique, and a supervised classifier based on {\em conditional random fields} to solve NER problems.
Knowledge-based techniques such 
as~\citename{zaghouani2010adapting}~\citeyear{zaghouani2010adapting} and~\citename{traboulsi2009arabic}~\citeyear{traboulsi2009arabic} propose local grammars with morphological stemming. % to perform NER. The work in~
\citename{ZaMaHaCicling2012Entity}~\citeyear{ZaMaHaCicling2012Entity} extract entities and events, and relations among them, from Arabic text using a hierarchy of manually built finite state machines driven by morphological features, 
and graph transformation algorithms.
Such techniques require advanced linguistic and programming expertise.

@@

{\bf Annotation tools.}
MMAX2 is a manual multi-level linguistic annotation tool with an XML 
based data model~\cite{mmax2}.
%It enables the user to create, browse, visualize, and query annotations and may be able to resolve coreference tags. 
BRAT~\cite{brat} and WordFreak~\cite{wordfreak} are manual 
multi-lingual user-friendly web-based annotators that allow the construction 
of entity and relation annotation corpora.
%They can be extended through plug-ins to enable automatic annotators and customized annotation visualization specifications.
Knowtator~\cite{ogren2006knowtator} is a general purpose incremental text annotation tool 
implemented as a Prot\'eg\'e~\cite{gennari2003evolution} plug-in. 
Prot\'eg\'e is an open-source platform with a suite of tools to construct domain models and knowledge-based applications with ontology. 
However, it doesn't support the Arabic language.

@@

\citename{kholidy2010towards}~\citeyear{kholidy2010towards} present an overview of annotation tools 
%and their Arabic-English word alignment issues 
and concludes with a set of rules and guidelines needed in an Arabic annotation alignment tool.
The work in~\citename{dukes2013supervised}~\citeyear{dukes2013supervised} presents a collaborative effort towards morphological and syntactic annotation of the Quran.
\citename{dorr2010interlingual}~\citeyear{dorr2010interlingual} present a framework for interlingual annotation of parallel text corpora with multi-level representations.
\citename{kulick2010consistent}~\citeyear{kulick2010consistent} presents the integration of the Standard Arabic Morphological Analyzer (SAMA) into the workflow of the Arabic Treebank.

The work in~\citename{smrz2004morphotrees}~\citeyear{smrz2004morphotrees} presents %the extension of TrEd, 
a customizable general purpose tree editor, with the Arabic MorphoTrees annotations. 
The MorphoTrees present the morphological analyses in a hierarchical organization based on common features.

Task specific annotation tools such as~\citename{alrahabi2006semantic}~\citeyear{alrahabi2006semantic} use enunciation semantic maps to automatically annotate directly reported Arabic and French speech.
AraTation is another task specific tool for semantic annotation of 
Arabic news using web ontology based semantic maps~\cite{saleh2009aratation}.
We differ in that \framework is general, and not task specific, and it uses 
morphology-based features as atomic terms.
Fassieh is a commercial Arabic text annotation tool that enables the production of large Arabic text corpora~\cite{attia2009fassieh}.
The tool supports Arabic text factorization including morphological analysis, POS tagging, 
full phonetic transcription, and lexical semantics analysis in an automatic mode.
Fassieh is not directly accessible to the research community and requires commercial licensing.
\framework is open source and differs in that it allows the user to build tag types and extract entities and relations from text.
